### 2021-11-11 analysis/03_experiment/rscripts/pre-processing.Rmd

---
title: "Pre-processing Data from Root Questions Corpus Study"
author: Morgan Moyer
date: March 22, 2021
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(lme4)
library(lmerTest)
library(multcomp) # not available for this version of R
library(tidyverse)
theme_set(theme_bw())
cbPalette <- c("#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73")
this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)
source("../../helpers.R")
```

```{r import data}
corp = read.table("../../../corpus/results/swbd.tab",sep="\t",header=T,quote="")
# Read the data into R.
d = read.csv("../data/exp03_main-merged.csv")
```

```{r combine raw data with corpus file}
# rename item_id from corpus file
names(corp)[names(corp) == "Item_ID"] <- "tgrep_id"
# filter from the database the tgrep_ids from the data
corp_match = corp %>%
  filter(tgrep_id %in% d$tgrep_id)

# join together
d <- left_join(d, corp_match, by="tgrep_id")

# take a look at the data
# summary(d)
# nrow(d)

# assign the negative values 0
d$rating[d$rating < 0] = 0
```

```{r take a look at the raw distribution, include = FALSE}
ggplot(d, aes(x=rating)) +
  geom_histogram()

summary(d) # this doesn't show any negative rating values....
tail(d)

# But clearly there are negative ratings....
negs = d %>%
  filter(rating < 0)
nrow(negs) # 140 negative ratings

```

# Native speakers
## We exclude any speakers who did not report English as a native language in the post-survey demographic questions.
```{r filter native English speakers, warning = FALSE}
nrow(d) # 174549
length(unique(d$workerid)) # 656
#filter out participants who did not report English as native language
d <- d %>%
  mutate(language = tolower(subject_information.language)) %>%
  filter(., grepl("eng", language))

# 25 participants removed for reporting non-english native language
nrow(d) # 107868
length(unique(d$workerid)) # 632
```

```{r comments and demographic information, eval=FALSE, include=FALSE}
length(unique(d$workerid)) # 656

d$time_in_minutes = as.numeric(as.character(d$time_in_minutes))
d$rating = as.numeric(d$rating)

length(unique(d$workerid)) #656

# look at comments
unique(d$subject_information.comments)

# fair price
ggplot(d, aes(x=subject_information.fairprice)) +
  geom_histogram(stat="count")

# overall assessment
ggplot(d, aes(x=subject_information.enjoyment)) +
  geom_histogram(stat="count")

# gender
ggplot(d, aes(x=subject_information.gender)) +
  geom_histogram(stat="count")

# education
ggplot(d, aes(x=subject_information.education)) +
  geom_histogram(stat="count")

# time_in_minutes
ggplot(d, aes(x=time_in_minutes)) +
  geom_histogram(stat="count")
mean(d$time_in_minutes)
```

```{r Look at practice trials, include= FALSE}
practice = d %>%
  filter(tgrep_id %in% c("example1", "example2", "example3", "example4"))

agr = practice %>%
  group_by(tgrep_id, paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

labels = c("Who can help spread the word?","Where can I get coffee around here?","Who came to the party?","How do I get to Central Park?")
names(labels) = c("example1","example2","example3", "example4")

ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position="dodge") +
  facet_wrap(~tgrep_id,labeller = labeller(tgrep_id = labels))
# ggsave("../graphs/main_practice_total.pdf")
# theme(axis.text.x = element_text(angle = 90))
```

```{r TODO: look at first response on practice trials only, eval=FALSE, include=FALSE}
prac_agr = practice %>%
  group_by(workerid,tgrep_id,paraphrase,rating) %>%
  summarise(count = n()) %>%
  group_by(workerid,tgrep_id) %>%
  mutate(total_per_ex = sum(count))
nrow(prac_agr) # 19566

prac_agr_rem = practice %>%
  group_by(workerid,tgrep_id,paraphrase,rating) %>%
  summarise(count = n()) %>%
  group_by(workerid,tgrep_id) %>%
  mutate(total_per_ex = sum(count)) %>%
  filter(total_per_ex > 4)
  # write.csv(.,"practice_to_edit.csv")
nrow(prac_agr_rem) # 2014

# fixed = read.csv("practice_edited.csv",header=TRUE)
nrow(fixed) # 68
head(fixed)
# remove that one column
# fixed = fixed[c(2:7)]

head(fixed)
prac_agr_keep = practice %>%
  group_by(workerid,tgrep_id,paraphrase,rating) %>%
  summarise(count = n()) %>%
  group_by(workerid,tgrep_id) %>%
  mutate(total_per_ex = sum(count)) %>%
  filter(total_per_ex <= 4)
nrow(prac_agr_keep) # 4088

practice_first = rbind(fixed,prac_agr_keep)
nrow(practice_first) #320
```

```{r Look at control trials}
controls = d %>%
  filter(grepl("control",tgrep_id))

# read in the file to have access to the items
cntrls = read.csv("../../../experiments/clean_corpus/controls.csv",header=TRUE,quote="")

# rename the item column in order to merge on it
names(cntrls)[names(cntrls) == "TGrepID"] <- "tgrep_id"

# join dfs together
c <- left_join(controls, cntrls, by="tgrep_id")
```

```{r Graph control trials, eval=FALSE, include=FALSE}
agr = c %>%
  group_by(EntireSentence,paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position="dodge") +
  facet_wrap(~EntireSentence, labeller = labeller(Sentence = label_wrap_gen(1)))
# ggsave("../graphs/main_controls.pdf")
```

```{r Remove subjects who failed 2/6 controls}
# for each control trial type, create a binary measure of whether the
# trial was passed
t = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="[_]") %>%
  group_by(workerid,paraphrase,trial) %>%
  filter(trial %in% c("movie", "book"), paraphrase %in% c("the")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
nrow(t) #1312
a1 = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="_") %>%
  group_by(workerid,paraphrase,trial) %>%
  filter(trial %in% c("novels", "cookies"), paraphrase %in% c("all")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
nrow(a1) #1312
a2 = c %>%
  separate(tgrep_id,into=c("tgrep_id","para","trial"),sep="_") %>%
  group_by(workerid,paraphrase,trial) %>% 
  filter(trial %in% c("tissue", "napkin"), paraphrase %in% c("a")) %>%
  mutate(control_passed = ifelse(rating > .5,"1","0"))
nrow(a2) #1312

# combine all those files together
con = rbind(t,a1,a2)
nrow(con) #3936

# filter out participants who failed more than 2 controls by taking the sum of 
# all the passed controls, and filtering out workerids who passed more than 2
failed_controls = con %>%
  filter(control_passed == "1") %>%
  group_by(workerid, control_passed) %>%
  summarise(sum_control_passed = n()) %>%
  filter(sum_control_passed < 4)

length(unique(failed_controls$workerid)) # 35
length(unique(failed_controls$workerid))/length(unique(d$workerid))*100 # 5.3%
```

# Focus on Test Items
```{r Look at raw test items}
# N before removing ppl who failed 2 controls 
length(unique(d$workerid)) #632

test = d %>%
  filter(!workerid %in% c(failed_controls$workerid)) %>%
  filter(!tgrep_id %in% c("example1", "example2", "example3", "example4","bot_check")) %>%
  filter(!grepl("control",tgrep_id))
# rename all to every
test$paraphrase[test$paraphrase == "all"] = "every"

# N after removing participants who failed 2/6 controls
length(unique(test$workerid)) #597

test[is.na(test)] <- 0
nrow(test) # 74860

agr = test %>%
  group_by(paraphrase) %>%
  mutate(mean_rating = mean(rating))

nrow(test) #71980
```

## Graph raw test items
```{r Graph raw test items, eval=FALSE, include=FALSE}
########################################################################
agr = test %>%
  group_by(paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position="dodge", show.legend = FALSE) +
  # ggtitle("Overall mean rating for each paraphrase") +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  theme(legend.position = "none") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("../graphs/test_overall.pdf")

agr = test %>%
  group_by(Wh,ModalPresent,paraphrase) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=ModalPresent)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=position_dodge(0.9)) +
  facet_wrap(~Wh) +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("../graphs/test_ModxWh.pdf")

```

# Remove Rhetorical Questions
```{r Remove rhetorical questions}
test_agr = test %>%
  group_by(tgrep_id, paraphrase) %>%
  summarize(mean_rating = mean(rating))

other_ratings = test %>%
  group_by(tgrep_id, paraphrase) %>%
  summarize(mean_rating = mean(rating)) %>%
  filter((mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="a"]) & 
           (mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="every"]) & 
           (mean_rating[paraphrase == "other"] > mean_rating[paraphrase=="the"]))

# How much data removed?
nrow(other_ratings)/nrow(test_agr)*100 # 15.8%
nrow(test_agr)#3980

or_ids = other_ratings$tgrep_id
test_other = test %>%
  filter(tgrep_id %in% or_ids)

# filter out those bad guys
test_norm = test %>%
  filter(!tgrep_id %in% or_ids)
nrow(test_norm)/nrow(test)*100 # 84.8%

nrow(test_norm) # 60744

# write.csv(test_norm,"../data/root_no_rhetorical.csv")
```

# Normalize ratings

```{r Normalize ratings to make a probability distribution}
# Remove 'other' ratings
critical = test_norm %>%
  filter(paraphrase %in% c("every","a","the"))

# unique by-item/by-participant combo
critical$ids = paste(critical$workerid,critical$tgrep_id)

critical$ModalPresent = as.factor(critical$ModalPresent)
critical$Wh = as.factor(critical$Wh)
critical$paraphrase = as.factor(critical$paraphrase)

# Determine for each observation (by-participant by-item), get the sum of the 
# 3 ratings
cr = critical %>%
  group_by(ids) %>%
  summarize(rating_sum = sum(rating))

# Join the dfs together
critical = critical %>%
  left_join(cr, by="ids")

# For each of the three paraphrase ratings, divide by sum of ratings for that item
critical$factors = paste(critical$ids,critical$paraphrase)
normed_agr = critical %>%
  group_by(factors) %>%
  summarise(normed_rating = rating/rating_sum) %>%
  drop_na() # this removes ALOT of rows

normed = merge(normed_agr,critical,by='factors')
normed[is.na(normed$ModalPresent)] <- "no"

# FIND OUT:
# are there particular items that show bimodality between "other" and another para?
# FIND OUT:
# Individual differences using KL divergence

length(unique(normed$tgrep_id)) #842

# subset the data to just the relevant columns so that we don't encounter \
# a large file storage problem on GitHub
normed <- normed[,c("workerid","tgrep_id","Sentence","paraphrase","ModalPresent","normed_rating","QuestionType","Finite","Modal","Wh")]

# save to .csv to load into analysis script
# write.csv(normed,"../data/normed.csv")

nrow(normed) # 41592
```

# Graph Normed Data

## Overall
```{r Graph normed data Overall}
agr = normed %>%
  group_by(paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
View(agr)

ggplot(agr,aes(x=paraphrase, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  # geom_errorbar(aes(ymin = YMin, ymax = YMax),width=.25, position=position_dodge(width = 0.9)) +
  # 
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position="dodge", show.legend = FALSE) +
  # ggtitle("Mean rating for 'a' vs. 'every'") +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  # ylim(0,.6) +
  theme(legend.position = "none") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("../graphs/overall.pdf",width=2.5,height=2)
```

## Modal x Wh-Word
```{r Graph Interaction between Mod and WH}
agr = normed %>%
  group_by(Wh,ModalPresent,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

# Re-Order the WH-levels by overall composition of DB
agr$Wh <- factor(agr$Wh, levels=c("what","how","where","why","who","when"))

ggplot(agr,aes(x=paraphrase, y=mean_rating, alpha=ModalPresent, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=.25,position=position_dodge(0.9)) +
  facet_wrap(~Wh, ncol=2) +
  xlab("Paraphrase") +
  ylab("Mean rating") +
  guides(fill=FALSE) +
  guides(alpha=guide_legend(title="Modal present")) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,-5,-5),legend.spacing.y = unit(0.001, 'cm')) +
    scale_fill_manual(values=cbPalette) +
    scale_color_manual(values=cbPalette) +
    scale_alpha_discrete(range = c(.5,1))

# ggsave("../graphs/modxwh_long.pdf",width=5,height=3)
```

## Wh-Word

```{r Graph Wh only}
agr = normed %>%
  group_by(paraphrase,Wh) %>%
  summarize(mean_rating = mean(rating), CILow = ci.low(rating), CIHigh = ci.high(rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()
# Re-Order the WH-levels by overall composition of DB
agr$Wh <- factor(agr$Wh, levels=c("what","how","where","why","who","when"))
ggplot(agr,aes(x=Wh, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=.25,position=position_dodge(0.9))  +
  # ggtitle("Mean rating for Wh-Word") +
  xlab("Wh-Word") +
  ylab("Mean rating") +
  theme(legend.title = element_blank()) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("../graphs/final_normed_wh.pdf")
```

## ModalPresent

```{r Graph ModalPresent only}
agr = normed %>%
  group_by(paraphrase,ModalPresent) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh) %>%
  drop_na()

ggplot(agr,aes(x=ModalPresent, y=mean_rating, fill=paraphrase)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=.25,position=position_dodge(0.9)) +
  xlab("Modal Present") +
  ylab("Mean rating") +
  theme(legend.title = element_blank()) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# legend.spacing.y = unit(-10, 'cm'))
# guides(fill=guide_legend(title="Paraphrase"))
# ggsave("../graphs/final_normed_modalpresent.pdf")
```

## Modals

```{r Look at all the modals}
mod = normed %>%
  filter(ModalPresent %in% c("yes")) %>%
  group_by(Modal,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), CILow = ci.low(normed_rating), CIHigh = ci.high(normed_rating)) %>%
  mutate(YMin = mean_rating - CILow, YMax = mean_rating + CIHigh)

ggplot(mod, aes(x=paraphrase,y=mean_rating,fill=paraphrase)) +
  geom_bar(stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=.25,position=position_dodge(0.9)) +
  facet_wrap(~Modal,ncol=3) +
  theme(legend.title = element_blank()) +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top", # c(.5,1)
        legend.direction = "horizontal") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("../graphs/final_normed_modals.pdf")
```

```{r Look at specific items, include=FALSE}

the_high = normed %>%
  filter((paraphrase == "the")) %>% #  & (Wh == "why")
  group_by(tgrep_id,Question) %>%
  summarize(mean_rating = mean(normed_rating), sd = sd(normed_rating)) %>%
  filter(mean_rating > .5)
# View(the_high)

a_high = normed %>%
  filter((paraphrase %in% c("a")) & (Wh == "who")) %>% #     & (ModalPresent == "no")
  group_by(tgrep_id,Question) %>%
  summarize(mean_rating = mean(normed_rating), sd = sd(normed_rating)) %>%
  filter(mean_rating > .3)
# View(a_high)

all_high = normed %>%
  filter((paraphrase %in% c("every")) & (Wh == "what")) %>% #  & (Wh == "when") & (ModalPresent == "yes")
  group_by(tgrep_id,Question) %>%
  summarize(mean_rating = mean(normed_rating), sd = sd(normed_rating)) %>%
  filter(mean_rating > .1)
# View(all_high)

# Look at all ratings for a specfic example
ex = normed %>%
  filter(tgrep_id %in% c("83839:35")) %>%
  group_by(paraphrase) %>%
  summarize(mean_rating = mean(normed_rating), sd = sd(normed_rating))
# View(ex)

```