### 2021-11-11 analysis/exploratory/all_context/rscripts/main_analysis.Rmd

---
title: "Analysis of ROOT & EMBEDDED QUESTIONS TOGETHER"
author: "mcmoyer"
date: "March 30, 2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
library(lme4)
library(lmerTest)
library(multcomp) # not available for this version of R
library(tidyverse)
theme_set(theme_bw())

this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)

cbPalette <- c("#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73")
```

########################################################################
# Read the data into R.
```{r}
rq = read.csv("../../../03_experiment/data/normed.csv")
eq = read.csv("../../../04_experiment/data/normed.csv")
rq = rq[,c("workerid","tgrep_id","ModalPresent","Modal","QuestionType","Question","Sentence","paraphrase","normed_rating","Finite","Wh")]
eq = eq[,c("workerid","tgrep_id","ModalPresent","Modal","QuestionType","Question","Sentence","paraphrase","normed_rating","Finite","Wh","VerbLemma")]

rq$QuestionType = "root"
rq$VerbLemma = ""

d = rbind(rq,eq)
table(d$QuestionType)

```

# Center and set contrast levels
```{r}
d$ModalPresent = as.factor(d$ModalPresent)
d$Wh = as.factor(d$Wh)
d$paraphrase = as.factor(d$paraphrase)
d$QuestionType = as.factor(d$QuestionType)

# set the reference levels
contrasts(d$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

# this won't work if the 'other' category is included, so REDO
contrasts(d$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))

centered = d
centered$cModalPresent = as.numeric(centered$ModalPresent) - mean(as.numeric(centered$ModalPresent))

```

# Full Model
```{r}
########################################################################
# model with random slopes no matrix verb predictor
# m.full = lmer(rating ~ cModalPresent*Wh*paraphrase + (1+paraphrase+Wh+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=centered,REML=FALSE) 
# summary(m.full)
# saveRDS(m.full, "../model-context-full.rds")

m.full <- readRDS("../model-context-full.rds")
summary(m.full)

```

# Break up models by Wh

## What
```{r}
########################################################################
# "what" model
d_what = centered %>% 
  filter(Wh=="what") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
contrasts(d_what$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_what$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.what = lmer(normed_rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|VerbLemma) + (1+paraphrase|tgrep_id), data=d_what)
# saveRDS(m.what, "../model-context-what-QT.rds")

m.what <- readRDS("../model-context-what-QT.rds")
summary(m.what)

# m.what.simple = lmer(normed_rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|VerbLemma)  + (1+paraphrase|tgrep_id), data=d_what) 
# summary(m.what.simple)
```

## How
```{r}
# "how" model
d_how = centered %>% 
  filter(Wh=="how") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
contrasts(d_how$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_how$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.how = lmer(normed_rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|VerbLemma)  + (1+paraphrase|tgrep_id), data=d_how)
# saveRDS(m.how, "../model-context-how-QT.rds")

m.how <- readRDS("../model-context-how-QT.rds")
summary(m.how)

# m.how.simple = lmer(normed_rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase|tgrep_id), data=d_how) 
# summary(m.how.simple)
```

## Where

```{r}
d_where = centered %>% 
  filter(Wh=="where") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
contrasts(d_where$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_where$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.where = lmer(normed_rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|VerbLemma) + (1+paraphrase|tgrep_id), data=d_where)
# saveRDS(m.where, "../model-context-where-QT.rds")

m.where <- readRDS("../model-context-where-QT.rds")
summary(m.where)

# m.where.simple = lmer(normed_rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_where) 
# summary(m.where.simple)
```

## Why

```{r}
d_why = centered %>% 
  filter(Wh=="why") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
contrasts(d_why$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_why$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.why = lmer(normed_rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|VerbLemma) + (1+paraphrase|tgrep_id), data=d_why) 
# saveRDS(m.why, "../model-context-why-QT.rds")

m.why <- readRDS("../model-context-why-QT.rds")
summary(m.why)
# m.why.simple = lmer(rating ~ paraphrase*cQuestionType*cModalPresent-cModalPresent + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_why) 
# summary(m.why.simple)
```

## Who

```{r}
d_who = centered %>% 
  filter(Wh=="who") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
contrasts(d_who$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_who$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.who = lmer(normed_rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase|workerid)+ (1+paraphrase|VerbLemma) + (1+paraphrase|tgrep_id), data=d_who) 
# saveRDS(m.who, "../model-context-who-QT.rds")

m.who <- readRDS("../model-context-who-QT.rds")
summary(m.who)

# m.who.simple = lmer(normed_rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase|workerid) + (1+paraphrase|tgrep_id), data=d_who) 
# summary(m.who.simple)
```

## When

```{r}
# have to use less random effects structure
d_when = centered %>% 
  filter(Wh=="when") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
contrasts(d_when$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_when$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.when = lmer(normed_rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase|VerbLemma) + (1+paraphrase|tgrep_id), data=d_when) 
# saveRDS(m.when, "../model-context-when-QT.rds")

m.when <- readRDS("../model-context-when-QT.rds")
summary(m.when)

# simple effects
# m.when.simple = lmer(normed_rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase|tgrep_id), data=d_when) 
# summary(m.when.simple)
# 
```