### 2021-11-11 analysis/exploratory/context_versus_nocontext/rscripts/kl_divergence.Rmd

---
title: "Exploratory analysis of context: KL Divergence"
author: Morgan Moyer
date: May 18, 2021
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
library(ggplot2)
library(lme4)
library(lmerTest)
library(multcomp) # not available for this version of R
library(philentropy)
library(stringr)
library(textstem)
library(tidyverse)
theme_set(theme_bw())
cbPalette <- c("#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73")
this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)
source("../../../helpers.R")
```

```{r read in all the data, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
context_root = read.csv("../../../03_experiment/data/normed.csv") 
context_root$Task <- "context"
context_root$VerbLemma <- ""
context_root = context_root[,!(names(context_root) %in% c("X"))]

context_emb = read.csv("../../../04_experiment/data/normed.csv") 
context_emb$Task <- "context"
context_emb = context_emb[,!(names(context_emb) %in% c("X"))]

nocontext_root = read.csv("../../../05_experiment/data/normed_nocontext_root.csv")
nocontext_root$VerbLemma <- ""
nocontext_root$Task <- "nocontext"
nocontext_root = nocontext_root[,!(names(nocontext_root) %in% c("X"))]

nocontext_emb = read.csv("../../../06_experiment/data/normed_nocontext_embedded.csv")
nocontext_emb$Task <- "nocontext"
nocontext_emb = nocontext_emb[,!(names(nocontext_emb) %in% c("X"))]
```

```{r join all the dfs together, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
context <- full_join(context_root,context_emb)
nocontext <- full_join(nocontext_root,nocontext_emb)
d <- full_join(context, nocontext)

# sanity check the join
nrow(context_root) + nrow(context_emb) + nrow(nocontext_root) + nrow(nocontext_emb)
nrow(d)
```

# Calculate KL-Divergence

KLD(P || Q) = sum x in X: (P(x) * log2(P(x)/Q(x)))

Compare to the uniform distribution.

```{r widen the data, warning=FALSE, echo=FALSE, message=FALSE}
# Aggregate ratings across subject and convert to wide
d.wide.context <- d %>%
  filter(Task == "context") %>%
  group_by(tgrep_id,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating)) %>%
  select(tgrep_id,paraphrase,mean_rating) %>%
  pivot_wider(names_from=paraphrase,values_from=mean_rating)

d.wide.nocontext <- d %>%
  filter(Task == "nocontext") %>%
  group_by(tgrep_id,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating)) %>%
  select(tgrep_id,paraphrase,mean_rating) %>%
  pivot_wider(names_from=paraphrase,values_from=mean_rating)
```

```{r calculate kld, warning=FALSE, echo=FALSE, message=FALSE}
kls.context <- d.wide.context %>%
  group_by(tgrep_id) %>%
  summarize(kld = KL(rbind(c(a,every,the),c(1/3,1/3,1/3)),unit="log2"))
kls.context$Task <- "context"

kls.nocontext <- d.wide.nocontext %>%
  group_by(tgrep_id) %>%
  summarize(kld = KL(rbind(c(a,every,the),c(1/3,1/3,1/3)),unit="log2"))
kls.nocontext$Task <- "nocontext"
```

```{r combine together, warning=FALSE, echo=FALSE, message=FALSE}
d.kls <- full_join(kls.nocontext,kls.context)
```

```{r factors, warning=FALSE, echo=FALSE, message=FALSE}
# look at linguistic factors
d.factors <- unique(d[,c("tgrep_id","ModalPresent","Wh","QuestionType")])
kls.factors <- left_join(d.kls,d.factors,by="tgrep_id")
```

## Plot the KLD between context and no context

### Overall
```{r, warning=FALSE, echo=FALSE, message=FALSE}
ggplot(data = d.kls, aes(x=Task, y = kld, fill=Task)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("graphs/denisty_kld_context.pdf",width=6,height=3)
```

### Factors
```{r, warning=FALSE, echo=FALSE, message=FALSE}
agr <- kls.factors %>% 
  group_by(Task, ModalPresent,Wh) %>% 
  summarize(mean_kld = mean(kld), 
            CILow = ci.low(kld), 
            CIHigh = ci.high(kld)) %>%
  mutate(YMin = mean_kld - CILow, 
         YMax = mean_kld + CIHigh) %>%
  drop_na()

# Re-Order the WH-levels by overall composition of DB
agr$Wh <- factor(agr$Wh, levels=c("what","how","where","why","who","when"))

ggplot(data = agr, aes(x = Task, y=mean_kld, fill=ModalPresent)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),
                width=.25,
                position="dodge", 
                show.legend = FALSE) +
  facet_wrap(~Wh,ncol=2) +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette) +
  scale_alpha_discrete(range = c(.5,1))
# ggsave("graphs/denisty_kld_context_factors.pdf",width=6,height=6)
```

## Difference in KLD by factors
```{r, warning=FALSE, echo=FALSE, message=FALSE}
kls.diffs.factors <- left_join(kls.difs.rel,d.factors,by="tgrep_id")

ggplot(data = kls.diffs.factors, aes(x = difference_in_kld, fill=ModalPresent)) +
  geom_density(alpha = .4) +
  facet_wrap(~Wh) +
  scale_fill_manual(values=cbPalette) +
  scale_color_manual(values=cbPalette)
# ggsave("graphs/difference_in_kld_factors_density.pdf",width=10,height=8)

#plot all the individuals mean chosen
ggplot(kls.diffs.factors, aes(x=Wh,y=difference_in_kld,fill=ModalPresent)) +
  geom_bar(stat="identity",position="dodge")
# ggsave("graphs/difference_in_kld_by-item_factors.pdf",width=10,height=8)

```

## Does Task predict KLD?

- Yes: going from Context to NoContext yields a decrease in KLD
- Interpretation: There is less information loss in the Context Task than in the NoContext Task
```{r, warning=FALSE, echo=FALSE, message=FALSE}
# just recode the factors so that negative values are nocontext and positive values are context
d.kls$Task <- as.factor(d.kls$Task)
d.kls$Task <- recode_factor(d.kls$Task, context=1, nocontext=0)

d.kls$tgrep_id = as.factor(d.kls$tgrep_id)
d.kls$cTask = as.numeric(as.factor(d.kls$Task)) - mean(as.numeric(as.factor(d.kls$Task)))

m.kld <- lmer(kld~cTask + (1|tgrep_id), data=d.kls)
summary(m.kld)

plot(m.kld)
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}

d.kls$Fitted = fitted(m.kld)
ggplot(d.kls, aes(x=Fitted,y=kld)) +
  geom_point() +
  geom_smooth(method="lm") +
  xlim(0,1) +
  ylim(0,1) +
  ylab("Empirical rating") +
  xlab("Predicted rating")
# ggsave("../graphs/model_fit_0.pdf",width=5,height=4)
cor(d.kls$Fitted,d.kls$kld)
```