### 2021-11-11 analysis/exploratory/question_type_context/main_analysis.Rmd

---
title: "Analysis of Root + Embedded Questions, NO CONTEXT"
author: "mcmoyer"
date: "March 30, 2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE}
library(lme4)
library(lmerTest)
library(multcomp) # not available for this version of R
library(tidyverse)
theme_set(theme_bw())

this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)
source("../../../helpers.R")

cbPalette <- c("#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73")
```

# Read Data in
```{r}
d = read.csv("../data/normed.csv")
```

# Center and set contrast levels
```{r}
d$ModalPresent = as.factor(d$ModalPresent)
d$Wh = as.factor(d $Wh)
d$paraphrase = as.factor(d$paraphrase)
d$QuestionType = as.factor(d$QuestionType)

# Need to first figure out which verbs to look at
d$VerbLemma = as.factor(d$VerbLemma)

# Set the reference levels
contrasts(d$ModalPresent) # once this is centered, the reference level won't matter anymore
contrasts(d$paraphrase)
contrasts(d$Wh)

contrasts(d$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

contrasts(d$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))

centered = cbind(d,myCenter(d["ModalPresent"]))

```

# Full Model
```{r}
# m.full = lmer(rating ~ cModalPresent*Wh*paraphrase + (1+paraphrase+Wh+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=centered,REML=FALSE)

# saveRDS(m.full, "model-nocontext-full.rds")
nocontext_model <- readRDS("model-nocontext-full.rds")
summary(nocontext_model)
```

# Models broken up by Wh

## What
```{r}
# "what" model
d_what = centered %>% 
  filter(Wh=="what") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
contrasts(d_what$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_what$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.what = lmer(rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_what)
# saveRDS(m.what, "model-nocontext-what-QT.rds")

m.what <- readRDS("model-nocontext-what-QT.rds")
summary(m.what)

# m.what.simple = lmer(rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_what) 
# summary(m.what.simple)
```

## How

```{r}
# "how" model
d_how = centered %>% 
  filter(Wh=="how") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  droplevels()
contrasts(d_how$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_how$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.how = lmer(rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_how)
# saveRDS(m.how, "model-nocontext-how.rds")

m.how <- readRDS("model-nocontext-how.rds")
summary(m.how)

# m.how.simple = lmer(rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase|tgrep_id), data=d_how) 
# summary(m.how.simple)
```

## Where
```{r}
# "where" model
d_where = d %>% 
  filter(Wh=="where") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  droplevels()
contrasts(d_where$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_where$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.where = lmer(rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_where)
# saveRDS(m.where, "model-nocontext-where.rds")

m.where <- readRDS("model-nocontext-where.rds")
summary(m.where)

# m.where.simple = lmer(rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_where) 
# summary(m.where.simple)
```

## Why

```{r}
d_why = d %>% 
  filter(Wh=="why") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  droplevels()
contrasts(d_why$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_why$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.why = lmer(rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_why)
# saveRDS(m.why, "model-nocontext-why.rds")

m.why <- readRDS("model-nocontext-why.rds")
summary(m.why)

# m.why.simple = lmer(rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_why) 
# summary(m.why.simple)
```

# Who
```{r}
d_who = d %>% 
  filter(Wh=="who") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  droplevels()
contrasts(d_who$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_who$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.who = lmer(rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase|workerid) + (1+paraphrase|tgrep_id), data=d_who)
# saveRDS(m.who, "model-nocontext-who.rds")

m.who <- readRDS("model-nocontext-who.rds")
summary(m.who)

# m.who.simple = lmer(rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase|workerid) + (1+paraphrase|tgrep_id), data=d_who) 
# summary(m.who.simple)
```

## When
```{r}
# have to use less random effects structure
d_when = d %>% 
  filter(Wh=="when") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  droplevels()
contrasts(d_when$paraphrase) = cbind("a.vs.every"=c(1,0,0),"the.vs.every"=c(0,0,1))
contrasts(d_when$QuestionType) = cbind("emb.vs.sai"=c(1,0,0),"root.vs.sai"=c(0,0,1))

# m.when = lmer(rating ~ cModalPresent*paraphrase*QuestionType + (1+paraphrase|tgrep_id), data=d_when)
# saveRDS(m.when, "model-nocontext-when.rds")

m.when <- readRDS("model-nocontext-when.rds")
summary(m.when)

# simple effects
# m.when.simple = lmer(rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase|tgrep_id), data=d_when) 
# summary(m.when.simple)
```