### 2021-11-29 analysis/exploratory/context_versus_nocontext/rscripts/examples.Rmd
updated talk for linguae

---
title: "Exploratory analysis of context: Examples"
author: Morgan Moyer
date: October 19, 2021
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
library(ggplot2)
library(lme4)
library(lmerTest)
library(multcomp) # not available for this version of R
library(philentropy)
library(stringr)
library(textstem)
library(tidyverse)
theme_set(theme_bw())
cbPalette <- c("#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73","#56B4E9", "#D55E00", "#009E73","#999999", "#E69F00","#009E73")
this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)
source("../../../helpers.R")
```

```{r read in all the data, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
context_root = read.csv("../../../03_experiment/data/normed.csv") 
context_root$Task <- "context"
context_root$VerbLemma <- ""
context_root = context_root[,!(names(context_root) %in% c("X"))]

context_emb = read.csv("../../../04_experiment/data/normed.csv") 
context_emb$Task <- "context"
context_emb = context_emb[,!(names(context_emb) %in% c("X"))]

nocontext_root = read.csv("../../../05_experiment/data/normed_nocontext_root.csv")
nocontext_root$VerbLemma <- ""
nocontext_root$Task <- "nocontext"
nocontext_root = nocontext_root[,!(names(nocontext_root) %in% c("X"))]

nocontext_emb = read.csv("../../../06_experiment/data/normed_nocontext_embedded.csv")
nocontext_emb$Task <- "nocontext"
nocontext_emb = nocontext_emb[,!(names(nocontext_emb) %in% c("X"))]
```

```{r join all the dfs together, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
context <- full_join(context_root,context_emb)
nocontext <- full_join(nocontext_root,nocontext_emb)
d <- full_join(context, nocontext)

# sanity check the join
nrow(context_root) + nrow(context_emb) + nrow(nocontext_root) + nrow(nocontext_emb)
nrow(d)
```

```{r bring in corpus to see the context, include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
corp = read.csv("../../../../corpus/analysis/swbd_contexts.csv")
names(corp)[names(corp) == "TGrepID"] <- "tgrep_id"
```

Interesting MA with domain restriction

```{r, warning=FALSE, echo=FALSE, message=FALSE}
corp %>%
  filter(tgrep_id == "2512:34") %>% 
  select(PreceedingContext) %>% 
  slice(1)
```

```{r , warning=FALSE, echo=FALSE, message=FALSE}
d %>%
  filter(tgrep_id == "2512:34") %>% 
  select(Sentence) %>% 
  slice(1)
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}
d %>%
  filter(tgrep_id == "2512:34") %>%
  group_by(Task,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating))
```

World knowledge
```{r, warning=FALSE, echo=FALSE, message=FALSE}
corp %>%
  filter(tgrep_id == "101503:16") %>% 
  select(PreceedingContext) %>% 
  slice(1)
```

```{r , warning=FALSE, echo=FALSE, message=FALSE}
d %>%
  filter(tgrep_id == "101503:16") %>% 
  select(Sentence) %>% 
  slice(1)
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}
d %>%
  filter(tgrep_id == "101503:16") %>%
  group_by(Task,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating))
```

MS

```{r, warning=FALSE, echo=FALSE, message=FALSE}
corp %>%
  filter(tgrep_id == "38489:4") %>% 
  select(PreceedingContext) %>% 
  slice(1)
```

```{r , warning=FALSE, echo=FALSE, message=FALSE}
d %>%
  filter(tgrep_id == "38489:4") %>% 
  select(Sentence) %>% 
  slice(1)
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}
d %>%
  filter(tgrep_id == "38489:4") %>%
  group_by(Task,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating))
```

hmmm
```{r, warning=FALSE, echo=FALSE, message=FALSE}
corp %>%
  filter(tgrep_id == "11532:7") %>% 
  select(PreceedingContext) %>% 
  slice(1)
```

```{r , warning=FALSE, echo=FALSE, message=FALSE}
d %>%
  filter(tgrep_id == "11532:7") %>% 
  select(Sentence) %>% 
  slice(1)
```

```{r, warning=FALSE, echo=FALSE, message=FALSE}
d %>%
  filter(tgrep_id == "39551:19") %>%
  group_by(Task,paraphrase) %>%
  summarize(mean_rating = mean(normed_rating))
```